MCC Informativeness Kids - Additional data

### 2019-01-22 papers/cogsci2019/CogSCi2019_new.Rmd
CogSci paper new template

abstract: 
    "Pragmatic inferences are an integral part of language comprehension and learning. To recover the intended meaning of an utterance, listeners need to balance and integrate different sources in information. In a series of experiments, we studied how listeners integrate general expectations about speakers with expectations that are specific to the interactional history with  a particular speaker. We used a Bayesian cognitive model to formalize the integration process. In Experiment 1 and 2 we replicated findings showing that listeners make inferences based on general and speaker specific expectations. Using the empirical measurements from these experiments, we generated model predictions about how they should be integrated. We tested these predictions in Experiment 3, which combined both expectations in a single design. Experiment 4 replicated and extended Experiment 3 to a broader set of conditions. Across both experiments, listeners based their inferences on both types of expectations. Overall, model predictions and data were highly correlated. Predictions from the model incorporating both types of information described listeners’ behavior better compared to alternative models based on only one type of information. This research shows that listeners flexibly integrate different forms of social expectations across a range of contexts, a process which can be described using Bayesian cognitive models. Our approach highlights how experiments and computational models can be used in conjunction to inform theories about language use and comprehension."

keywords:
    "Pragmatics; Word learning; Common ground; Bayesian models"

output: cogsci2016::cogsci_paper
---

>>>>

abstract: >
    Pragmatic inferences are an integral part of language comprehension and learning. To recover the intended meaning of an utterance, listeners need to balance and integrate different sources in information. In a series of experiments, we studied how listeners integrate general expectations about speakers with expectations that are specific to the interactional history with  a particular speaker. We used a Bayesian cognitive model to formalize the integration process. In Experiment 1 and 2 we replicated findings showing that listeners make inferences based on general and speaker specific expectations. Using the empirical measurements from these experiments, we generated model predictions about how they should be integrated. We tested these predictions in Experiment 3, which combined both expectations in a single design. Experiment 4 replicated and extended Experiment 3 to a broader set of conditions. Across both experiments, listeners based their inferences on both types of expectations. Overall, model predictions and data were highly correlated. Predictions from the model incorporating both types of information described listeners’ behavior better compared to alternative models based on only one type of information. This research shows that listeners flexibly integrate different forms of social expectations across a range of contexts, a process which can be described using Bayesian cognitive models. Our approach highlights how experiments and computational models can be used in conjunction to inform theories about language use and comprehension.

keywords: >
    Pragmatics; Word learning; Common ground; Bayesian models

output: cogsci2016::cogsci_paper
#final-submission: \cogscifinalcopy
---

### 2019-01-22 papers/cogsci2019/CogSCi2019_new.tex
CogSci paper new template

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

@@

\title{Integrating Common Ground and Informativeness in Pragmatic Word Learning}

@@

In our model, we used the results from Experiment 1 and 2 to specify
\(\alpha\) as well as \(P_S(r_s)\). The former was set to yield the
proportion measured in Experiment 1 in a model with uniform priors. The
prior probability for each object was the proportion with which this
object was chosen in Experiment
2\footnote{Proportions were measured when participants chose between two objects. However, in Experiment 3, three objects were involved. For each object we used the proportion measured in Experiment 2 as the prior probability.This approached spread out the absolute probability mass for each object but conserved the relative relation between objects.}.
Based on these parameter settings, we predicted the proportion with
which listeners will choose the less frequent object in each unique
combination of alignment and speaker identity separately for each prior
manipulation (eight conditions total, see Figure \ref{fig:plotexp3}). We
compared the fit of this pragmatic model to two alternative models using
Bayes Factors (BF). The first alternative model ignored the speaker
specific expectations (flat prior model) while the second ignored the
informativeness inference (prior only model). To capture the idea that
behavioral data is to some degree noisy, all models included a ``noise
parameter'', which reflects that, with a certain probability,
participants respond randomly instead of in line with the intended
manipulation on a given trial. Noise parameters are estimated based on
the data.

@@

>>>>

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{apacite}

% KM added 1/4/18 to allow control of blind submission

@@

\title{How to Make a Proceedings Paper Submission}

@@

In our model, we used the results from Experiment 1 and 2 to specify
\(\alpha\) as well as \(P_S(r_s)\). The former was set to yield the
proportion measured in Experiment 1 in a model with uniform priors. The
prior probability for each object was the proportion with which this
object was chosen in Experiment
2\footnote{Proportions were measured when participants chose between two objects. However, in Experiment 3, three objects were involved. For each object we used the proportion measured in Experiment 2 as the prior probability.This approached spread out the absolute probability mass for each object but conserved the relative relation between objects.}.
Based on these parameter settings, we predicted the proportion with
which listeners will choose the less frequent object in each unique
combination of alignment and speaker identity separately for each prior
manipulation (eight conditions total, see Figure \ref{fig:plotexp3}). We
compared the fit of this pragmatic model to two alternative models using
Bayes Factors (BF). The first alternative model ignored the speaker
specific expectations (flat prior model) while the second ignored the
informativeness inference (prior only model). To capture the idea that
behavioral data is to some degree noisy, all models included a noise
parameter, which reflects that, with a certain probability, participants
respond randomly instead of in line with the intended manipulation on a
given trial. Noise parameters are estimated based on the data.

@@

\bibliographystyle{apacite}

### 2019-01-22 stats/kids/kids_Ex3_pref.Rmd
CogSci paper new template

---
title: "MCC Ex3 Preference Kids"
author: "Manuel Bohn"
output: html_document
---

```{r import data, include = F}
###
library(tidyverse)
library(knitr)
library(langcog)
library(ggthemes)
library(jsonlite)
library(readxl)

@@

raw_data <- raw_data %>%
  mutate(subid = ifelse(subid == "181128_3_ex3p....","181128_3_ex3p" , subid),
         subid = ifelse(subid == "181128_1_ex3p.","181128_1_ex3p" , subid),
         subid = ifelse(subid == " 181218_10_ex3p","181218_10_ex3p" , subid))

raw_data%>%
  #group_by(paste(subid,test_date))%>%
  group_by(subid)%>%
  summarise(n= n())

unique(raw_data$subid)

@@

unique(d$check_age)

@@

# Sample size per age bin

```{r kids per age group,echo = F }

d %>%
  group_by(subage,subid)%>%
  summarise(n_trials = n())%>%
  mutate(full_data = ifelse(n_trials == 10,"yes","no"))%>%
  group_by(subage)%>%
  summarise(n = length(unique(subid)),
            n_full_data = sum(full_data == "yes"))%>%
  knitr::kable(digits = 2)

@@

# Datapoints per cell

```{r kids per age group,echo = F }

d %>%
  filter(trial != "filler1",trial != "filler2")%>%
  group_by(subage,speaker,alignment)%>%
  summarise(datapoints = n())%>%
  knitr::kable(digits = 2)

# Plot filler by age

```{r plot by age, echo= FALSE}
pf1 <- d %>%
      filter(trial == "filler1" | trial == "filler2")%>%
  group_by(subage, subid) %>%
  summarise(correct = mean(correct_inf))

@@

ggplot() +
  geom_jitter(data = pf1, aes(x = subage, y = correct, col = subage, alpha = .2),width = .1,height = .025)+
  geom_pointrange(data = pf2, aes(x = subage, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())
```

# Plot by age group

```{r plot by age, echo= FALSE}

@@

ggplot() +
  geom_jitter(data = p1, aes(x = alignment, y = correct, col = alignment, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p2, aes(x = alignment, y = mean, col = alignment, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_grid(speaker~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_colour_solarized(name="Control",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

# Plot by age continous

```{r plot by age, echo= FALSE}

p3 <- d %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  group_by(speaker, alignment ,age_num, subid) %>%
  summarise(correct = mean(correct_inf))

ggplot(data = p3, aes(x = age_num, y = correct_inf)) +
  geom_jitter(width = .025,height = .025, alpha = .2)+
  geom_smooth(method = "glm", method.args = list(family = "binomial"),  se = T,  alpha = .5, size = .4)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="Age",y="Proportion Expected Choice")+
  theme_few() +
  facet_grid(speaker~alignment)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+
  coord_fixed(ratio = 2)

@@

```{r model}

@@

summary(mex3)
```

```{r chance_comp}
mod <- d %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  group_by(subage,speaker,alignment) %>%
  do(models = glmer(correct_inf ~ 1+ (1|subid),data=., family=binomial, control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))))

tidy(mod, models) %>%
  filter(term =="(Intercept)")
```

>>>>

---
title: "MCC Ex3 Preference Kids"
author: "Manuel Bohn"
output:
  html_document:
    code_folding: hide
---

```{r import data, include = F}
###
library(tidyverse)
library(knitr)
library(langcog)
library(ggthemes)
library(jsonlite)
library(readxl)
library(coda)

estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}

hdi_upper<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}

hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}

@@

raw_data <- raw_data %>%
  mutate(subid = ifelse(subid == "181128_3_ex3p....","181128_3_ex3p" , subid),
         subid = ifelse(subid == "181128_1_ex3p.","181128_1_ex3p" , subid),
         subid = ifelse(subid == " 181218_10_ex3p","181218_10_ex3p" , subid))
#unique(raw_data$subid)

@@

@@

## Overview

This is what the study looked like. The animal first commented on an object on one side, then on an object on the other side (preference manipulation) and then disappeared. Then either the same or a different animal returned (speaker change manipulation), turned to the tree with the two objects and said it had a e.g. "dax". Then, two new trees appeared, each with one type of object and children were asked to pick the tree with the "dax".

```{r setup, echo = F, fig.cap="Screenshot from the experiment", fig.align = "center"}

knitr::include_graphics("../../papers/figures/kids_ex3_setup.png")

@@

Manipulations:

* Age: Measured continuous 
* Speaker: same or different
* Alignment: congruent(preferred object = less frequent) or in-congruent (preferred = more frequent)

Crossing this results in four conditions (cells) per age group. Data was collected within subjects, with two (one if they decided to opt out earlier) trials per cell (combination of alignment and speaker).

Total N = `r length(unique(d$subid))`. We amended the preregistration saying that we want 80 data points per cell. The table below shows that we got this.

```{r kids per age group, echo = F }
d %>%
  filter(trial != "filler1",trial != "filler2")%>%
  group_by(subage,speaker,alignment)%>%
  summarise(n = length(unique(subid)),
            datapoints = n())%>%
  knitr::kable(digits = 2)
```

### Filler trials

Filler trials by age. Each child started with two filler trials. In these trials, each tree only had one feature. Children had no problem learning words in unambiguous cases, showing they get the basic procedure.

```{r filler by age, echo= FALSE, fig.height=3,fig.width=3}
pf1 <- d %>%
      filter(trial == "filler1" | trial == "filler2")%>%
  group_by(subage, subid) %>%
  summarise(correct = mean(correct_inf))

@@

ggplot() +
  geom_jitter(data = pf1, aes(x = subage, y = correct,  alpha = .2),width = .1,height = .025)+
  geom_pointrange(data = pf2, aes(x = subage, y = mean, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)
```

## Comparison to chance by age bin

### Plot

For the plot, data is aggregated within individual and age bin. Looking at the data this way, it looks like 3-year-olds do not care about speaker and alignment but have a slight overall tendency to pick the more informative object. 4-year-olds, on the other hand, seem to behave quite differently in the different cells.
```{r plot by age binned}

@@

ggplot() +
  geom_jitter(data = p1, aes(x = alignment, y = correct, col = alignment, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p2, aes(x = alignment, y = mean, col = alignment, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_grid(speaker~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_colour_solarized(name="Alignment")
```

### Model within each cell

Given the unequal number of trials per subject, we pre-registered models for the comparison to chance within each cell, testing the intercept against 0. Model structure was: `correct_inf ~ 1 + (1|subid) + (1|agent)`. Results are in line with what you see in the plot above.

```{r chance_comp, warning=F, message=F}
library(lme4)

mod <- d %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  group_by(subage,speaker,alignment) %>%
  do(models = glmer(correct_inf ~ 1+ (1|subid) + (1|agent),data=., family=binomial, control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))))

tidy(mod, models) %>%
  filter(term =="(Intercept)") %>%
  select(-term, -group)%>%
  kable(digits = 2)
```

## Plot by age continous

Here is the data plotted with age continuous. Regression lines and CIs are based on loess fits.

```{r plot by age}

p3 <- d %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  group_by(speaker, alignment ,age_num, subid)

ggplot(data = p3, aes(x = age_num, y = correct_inf)) +
  geom_jitter(width = .025,height = .025, alpha = .2)+
  geom_smooth(method = "loess",  se = T,  alpha = .5, size = .4, span = 2)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="Age",y="Proportion more informative")+
  theme_few() +
  facet_grid(speaker~alignment)+
  ylim(-0.05,1.05)+
  guides(alpha = F)

@@

## GLMM

The goal of the frequentest model was to see if the three manipulations (age, speaker and alignment have an effect). The model converges as pre-registered: `correct_inf ~ age_num*speaker*alignment + (speaker*alignment | subid) + (age_num*speaker*alignment | agent)`.

The significant 3-way interaction shows that all three components matter. This is good news, showing that all of the things we hypothesized to be important actually affected children's behavior.

```{r model, cache = T, warning=F, message=F}

@@

tidy(mex3) %>%
  filter(group =="fixed") %>%
  select(-group)%>%
  kable(digits = 2)

## Comparing data to model

### Models vs data by age contious

The black line is the loess fit based on the data (same as above). Models are shown in different colors. It looks like the RSA model is making the best predictions (relative to the others). The fit is far from perfect, but I guess that was expected. We will eventually compare models using Bayes Factors.

```{r comparison plot by age,warning=F, message=F, fig.height= 6, fig.width=10}

p_data <- d %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  mutate(age_bin = age_num - 3,
         Speaker = ifelse(speaker == "same", "same_speaker", "different_speaker"),
         Alignment = alignment)%>%
  group_by(Speaker, alignment ,age_bin, subid)

kids_model_comp_mean_ci <- readRDS("../saves/kids_model_comp_mean_ci.rds")

ggplot(kids_model_comp_mean_ci)+
  geom_ribbon(aes(x =age_bin, ymin = lci, ymax = uci, fill = model), alpha = .4) +
  geom_line(aes(age_bin, mean, col = model), size = 1, alpha = .5, linetype = 2) + 
  geom_smooth(data = p_data, aes(x = age_bin, y = correct_inf), col = "black", size = 0.8, method = "glm", method.args = list(family = "binomial"), se = T,  alpha = .7)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(y="proportion more informative")+
  scale_x_continuous(limits=c(0,2), name="age", breaks = c(0,0.5,1,1.5,2), labels = c(3.0,3.5,4.0,4.5,5)) +
  ylim(-0.05,1.05)+
  facet_grid(Speaker~Alignment)+
  theme_few(base_size = 20)+
  scale_color_viridis_d(direction = -1)+
  scale_fill_viridis_d(direction = -1)
```

```{r, include = F}
ggsave("../graphs/kids_all_model.pdf", width = 11.1, height = 8, scale = 1)
```

### Correlation plot by age bin

Not sure if this makes much sense: I binned and averaged the model predictions and the data by age to get correlation plots. Even though the correlation is higher for the prior only model, the predictions are systematically off.
```{r, cache = T, warning=F, message=F}

data <- d %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  mutate(Speaker = ifelse(speaker == "same", "same_speaker", "different_speaker"),
         Alignment = alignment)%>%
  group_by(Speaker, Alignment ,subage) %>%
  summarize(k = sum(correct_inf), n = n())%>%
  ungroup() %>%
  mutate(a = 1 + k,
         b = 1 + n - k,
         Data_ci_lower  = qbeta(.025, a, b),
         Data_ci_upper = qbeta(.975, a, b),
         Data = (a-1)/(a+b-2))%>%
  select(-a,-b,-n,-k)

kids_model_comp <- readRDS("../saves/kids_model_comp.rds")

model <- kids_model_comp %>%
  mutate(subage = ifelse(age_bin<1,"3","4"))%>%
  group_by(model,Speaker,Alignment,subage)%>%
  summarise(Model = mean(prop_informative),
            Model_ci_lower = hdi_lower(prop_informative),
            Model_ci_upper = hdi_upper(prop_informative))

cor_plot <- model %>%
  ungroup()%>%
  mutate(Data = rep(data$Data,3),
         Data_ci_lower = rep(data$Data_ci_lower,3),
         Data_ci_upper  =rep(data$Data_ci_upper,3))

library(ggpubr)

ggplot(data = cor_plot,aes(x = Model, y = Data, col = subage)) +
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 1, size = .5)+
  geom_point(size = 3)+
  geom_errorbar(aes(ymin = Data_ci_lower, ymax = Data_ci_upper),width = 0,size = .5)+
  geom_errorbarh(aes(xmin = Model_ci_lower, xmax = Model_ci_upper), height = 0,size = .5)+
  coord_fixed()+
  stat_cor(method = "pearson", label.x = 0.1, label.y = 0.9, aes(x = Model, y = Data), inherit.aes = F)+
  facet_grid(~model)+
  xlim(0,1)+ylim(0,1)+
  xlab("Model")+
  theme_few(base_size = 10) + 
  scale_colour_solarized()
```

### 2019-01-22 stats/kids/kids_inf_2.Rmd
CogSci paper new template

---
title: "MCC Informativeness Kids 2"
author: "Manuel Bohn"
output: html_document
---

@@

## importing data
files <- dir("~/Work/MCC/git-mcc/kids_info_data/")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_info_data/",f,sep="")
  jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
  date <- str_sub(jf,45,str_length(jf)-11)
  id <- as_data_frame(jd$data$data) %>% mutate(test_date = date)
  raw_data <- bind_rows(raw_data, id)
}

d_raw <-raw_data %>%
  mutate(test_date= ifelse(test_date == "18-10-10", paste(20,test_date,sep = ""),test_date),
         subid= ifelse(subid == "101018_8_inf", "181010_8_inf",subid),
         subid= ifelse(subid == "181031_3", "181031_3_inf",subid),
         test_date= ifelse(test_date == "18-XX-XX", "2018-10-10",test_date),
         test_date= ifelse(test_date == "18-09-12", "2018-09-12",test_date),
         experiment = "mcc_kids_inf_2")

log <- read_excel("../../MCC-subject_log.xlsx", 1)%>%
  filter(Condition == "inf")%>%
  select(subid,experimenter,keep_drop,sex,dob)

@@

write_csv(d, "data/kids_inf_data.csv")

>>>>

---
title: "MCC Informativeness Kids - Additional data"
author: "Manuel Bohn"
output: html_document
---

@@

## importing data
files <- dir("~/Work/MCC/git-mcc/kids_info_data_2/")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_info_data_2/",f,sep="")
  jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
  date <- str_sub(jf,47,str_length(jf)-11)
  id <- as_data_frame(jd$data$data) %>% mutate(test_date = date)
  raw_data <- bind_rows(raw_data, id)
}

length(unique(raw_data$subid))

d_raw <-raw_data %>%
  mutate(test_date = as.Date(test_date),
         experiment = "mcc_kids_inf_2_2")

log <- read_excel("../../../MCC-subject_log.xlsx", 1)%>%
  filter(Condition == "inf2")%>%
  select(subid,experimenter,keep_drop,sex,dob)

@@

length(unique(d$subid))

write_csv(d, "data/kids_inf_2_data.csv")
